---
title: "Hydraulic Fracturing and Induced Seismicity in Texas"
subtitle: "A Spatio-Temporal Analysis of Fracturing Activity and Microseismic Events (2010–2014)"
author: dilo
date: today

format:
  html:
    theme: yeti
    toc: true
    toc-depth: 3
    toc-location: right
    code-fold: true
    page-layout: full

bibliography: fracref.bib
link-citations: true
---
# Background and Motivation
Fossil fuels still dominate global energy systems, accounting for roughly 80% of global energy consumption and more than 80% of U.S. energy use [@css2025facts]. Over the past two decades, advances in horizontal drilling and hydraulic fracturing have enabled the large-scale development of unconventional oil and gas (UOG) resources trapped in low-permeability shale formations [@wang2015review]. These technologies inject high-pressure fluid (mostly water mixed with sand and other additives) into rock formations to create fractures that allow hydrocarbons to flow to the wellbore [@schultz2020hydra]. The result: from the early 2000s to 2012, U.S. gas production rose exponentially (<1% to 40%) [@nicot2014source].

Environmental and geophysical concerns have risen accordingly. Fracking operations typically require millions of gallons of water per well [@jackson2015depths] and generate large volumes of brine that must be managed/disposed of, often through underground disposal wells [@gregory2011water; @chen2016water]. Changes in subsurface fluid pressure associated with these processes can alter stress conditions along existing faults, potentially triggering induced seismicity [@ellsworth2013inject]. Aquifer contamination is another major concern, especially because hydraulic fracturing activities are exempt from SDWA regulation (*thanks Dick Cheney*) [@nicot2014source; @underhill2023outcome].

Since roughly 2008, earthquake activity has increased across parts of the central United States, with many events linked to wastewater disposal and other oil and gas activities [@walter2018natural]. In response to increasing seismicity and the need for improved monitoring (in other words it got so bad that the Texas legislature had to do something about it), Texas established the Texas Seismological Network (TexNet) to expand monitoring capabilities across the state [@savvaidis2019texnet]. Recent studies suggest that seismicity, particularly the Permian Basin, has increased dramatically in recent decades and is often associated with fluid injection at depth @skoumal2021prolif.

This project explores the spatial and temporal relationship between UOG development and earthquake activity in Texas using FracFocus well data, USGS earthquake catalogs, and TexNet seismic records [@fracfocus_data; @usgs_texas_earthquakes; @texnet_data_portal].

# Methodology -- *Big Oil, Big Data*
This project leverages multiple datasets describing both oil and gas activity and earthquake events in Texas, including:

Hydraulic fracturing disclosures from the FracFocus Chemical Disclosure Registry [@fracfocus_data]

Earthquake catalogs from the U.S. Geological Survey (USGS) [@usgs_texas_earthquakes]

Seismic monitoring data from the Texas Seismological Network (TexNet) [@texnet_data_portal]

FracFocus contains well-level records describing hydraulic fracturing operations, including injection volumes, well identifiers, location, chemical disclosures, and job start/end dates. Because the raw FracFocus dataset is highly untidy (understatement) and often contains multiple entries for a single operation, records were cleaned to sensible ranges, and aggregated by API well number and job end date. For each well, the final dataset retains total base water volume, total base non-water volume, total vertical depth, API number, as well as positioning and temporal data. I also construct a total volume measure to serve as a proxy for the intensity of a fracking job. The raw data is 7,060,962 rows (many rows per well for chemical disclosure purposes), after collapsing, filtering, and cleaning, I was left with 95,197 Texas wells from 9/5/2010 to 12/31/2024.

Earthquake data were obtained from the USGS Comprehensive Earthquake Catalog and supplemented with data from TexNet, a more thorough monitoring network [@savvaidis2019texnet]. There's a considerable difference in frequency prior to 2017 (not because there were less earthquakes, but because monitoring was scarce). The catalogs start to track closely around 2020/2021, I assume the USGS integrated the TexNet stations around then. For the sake of this project, I use USGS events prior to 2017, and TexNet events from 2017 onward.

After cleaning and standardizing these datasets, the analysis compares the spatial and temporal patterns of UOG operations and seismic events. Time-series plots and spatial maps are used explore these relationships, and well characteristics such as depth and injection volume will hopefully reveal some interesting patterns.

## Cleaning the frac focus data
The FracFocus data is quite rich, but also horribly messy and large. Narrowing an area of interest cuts it down a bit; this can be done by filtering for the first two digits (42 for TX) of a well's American Petroleum Institute identifier (`APINumber` column), or filtering in `StateName` as a backup. This still wasn't perfect, so I clipped the wells that only fall within the Texas borders sf object. The data, by nature, is also very long, as it is intended to disclose chemicals and quantities put in the ground at different points in a well's operation span (though somehow it's legal to say *"trade secret"* or *"proprietary blend"*). For an analysis concerned with the chemical disclosures, good luck, but because I'm only concerned with well position and final injection volume, so I collapsed by `APINumber`, `JobEndDate`, and took the maxes of `TotalBaseWaterVolume`, `TotalBaseNonWaterVolume`, and `TVD` (depth) so that I'd hopefully be left with unique wells, their position, and an idea of final injection volume. That condensed the FracFocus data considerably, but there were still issues (year 3012, impossible depth and volume values), so I handled those by applying sensible mins and maxes, and trimming observations outside the 99th percentile. There's better/different ways to handle the FracFocus data (see [@underhill2023outcome; @chen2016water]).

```{r eval= FALSE, echo= TRUE}
library(data.table)
library(dplyr)
library(lubridate)

files <- paste0('Data/FracFocusCSV/FracFocusRegistry_', 1:15, '.csv')

clean_and_collapse_frac <- function(file_path) {
  # function: reads a raw FracFocus CSV, filters for Texas wells, standardizes data types, 
  #     collapses the data by API Number and Job End Date, and removes impossible values
  # argument: file_path,  the file path to a fracfocus csv
  # output: a (hopefully) cleaned data frame of texas fracking well data
  
  df <- fread(file_path)
  df_clean <- df %>%
    mutate(
      APINumber = as.character(APINumber),
      StateName = toupper(StateName)
    ) %>%
    # filter using texas API numbers or state names incase there's errors
    filter(substr(APINumber, 1, 2) == '42' | StateName %in% c('TEXAS', 'TX')) %>%
    mutate(
      JobEndDate = as.Date(parse_date_time(JobEndDate, orders = c("mdy IM p", "mdy HMS", "ymd HMS", "mdy", "ymd"))), #makes sure the data formats are handled correctly
      TotalBaseWaterVolume = as.numeric(TotalBaseWaterVolume),
      TotalBaseNonWaterVolume = as.numeric(TotalBaseNonWaterVolume),
      TVD = as.numeric(TVD)
    ) %>%
    #collapse by API number and JobEndDate (to capture final injection volume)
    group_by(APINumber, JobEndDate) %>%
    summarize(
      Latitude = first(Latitude),
      Longitude = first(Longitude),
      TVD = max(TVD, na.rm = TRUE),
      TBWV = max(TotalBaseWaterVolume, na.rm = TRUE),
      TBNWV = max(TotalBaseNonWaterVolume, na.rm = TRUE),
      .groups = 'drop'
    ) %>%
    # remove the inf values that max() created for NA groups
    mutate(
      TVD = ifelse(is.infinite(TVD), NA, TVD),
      TBWV = ifelse(is.infinite(TBWV), NA, TBWV),
      TBNWV = ifelse(is.infinite(TBNWV), NA, TBNWV)
    ) %>%
    # create a total fluid volume column
    mutate(
      Total_Fluid_Volume = coalesce(TBWV, 0) + coalesce(TBNWV, 0),
      Total_Fluid_Volume = ifelse(is.na(TBWV) & is.na(TBNWV), NA, Total_Fluid_Volume)
    ) %>%
    # drop absurd values
    filter(
      TVD > 500 & TVD <= 25000, 
      Total_Fluid_Volume > 0 & Total_Fluid_Volume <= 100000000,
      # drop absurd dates (somebody wrote year 3012 in the raw data)
      JobEndDate >= as.Date('2010-01-01') & JobEndDate <= as.Date('2024-12-31')
    )
  return(df_clean)
}

# apply and bind
master_frac <- lapply(files, clean_and_collapse_frac) %>%
  bind_rows() %>%
  distinct(APINumber, JobEndDate, .keep_all = TRUE) 

# it still requires more cleaning, API and State Name didn't work perfectly
master_frac_patched <- master_frac %>%
  filter(
    # drop wells that aren't in texas
    Latitude >= 25.8 & Latitude <= 36.5,
    Longitude >= -106.6 & Longitude <= -93.5,
    # should be positive I think
    (TBNWV >= 0 | is.na(TBNWV))
  )

summary(master_frac_patched %>% select(Latitude, Longitude, TBNWV))
```

```{r eval= FALSE, echo= TRUE}
# hopefully the last bit of cleaning, because there's still some questionable values
tbwv_ceiling <- quantile(master_frac$TBWV, 0.99, na.rm= TRUE)
tbnwv_ceiling <- quantile(master_frac$TBNWV, 0.99, na.rm= TRUE)

master_frac_cleanest <- master_frac_patched %>%
  filter(
    TBWV <= tbwv_ceiling | is.na(TBWV),
    TBNWV <= tbnwv_ceiling | is.na(TBNWV),
    Total_Fluid_Volume > 0 )

summary(master_frac_cleanest %>% 
          select(TBWV, TBNWV, Total_Fluid_Volume))

# last thing since there's still some wells outside of TX

tx_boundary <- states(cb = TRUE, class = "sf") %>%
  filter(STUSPS == "TX") %>%
  st_transform(4326)

# convert wells to sf
wells_sf <- st_as_sf(
  master_frac_cleanest, 
  coords = c("Longitude", "Latitude"), 
  crs = 4326,
  remove = FALSE 
)

wells_sf <- st_filter(wells_sf, tx_boundary)

master_frac_cleanest <- st_drop_geometry(wells_sf)

fwrite(master_frac_cleanest, "Data/master_frac_cleanest.csv") 
```

## Scraping the USGS data
This is like what we did in class, but it gets 15 years of low-magnitude earthquake data. A `for` loop builds a url query for each year, adds that csv to an empty list, then binds those into the earthquake data for this project. Spatial bounding (a box around texas), pausing bewteen requests (`Sys.sleep(1.5)`), and `GET()` error handling from the `httr` package ensure this script doesn't blow up.
```{r eval= FALSE, echo= TRUE}
library(httr)
library(readr)
library(dplyr)
library(lubridate)

# make a box around texas
min_lat <- 25.8
max_lat <- 36.5
min_lon <- -106.6
max_lon <- -93.5

# 2010 to 2024 is my time of interest
years <- 2010:2025

# empty list for each year's data
usgs_list <- list()

# loop to query the usgs API
for(y in years) {
  start_time <- paste0(y, "-01-01")
  end_time   <- paste0(y, "-12-31")
  # build the url
  url <- paste0(
    "https://earthquake.usgs.gov/fdsnws/event/1/query?",
    "format=csv",
    "&starttime=", start_time,
    "&endtime=", end_time,
    "&minlatitude=", min_lat,
    "&maxlatitude=", max_lat,
    "&minlongitude=", min_lon,
    "&maxlongitude=", max_lon
    # there is no minmagnitude bexause I want small earthquakes
  )
  Sys.sleep(1.5)
  #this handles errors
  resp <- GET(url)
  if(status_code(resp) == 200) {
    yr_data <- read_csv(content(resp, "text", encoding = "UTF-8"), show_col_types = FALSE)
    usgs_list[[as.character(y)]] <- yr_data
  } else {
    warning(paste("Failed to fetch data for", y, "- Status Code:", status_code(resp)))
  }
}

# bind usgs years together
usgs_raw <- bind_rows(usgs_list)

# reassign cols to match texnet
usgs_clean <- usgs_raw %>%
  select(
    Time = time,
    Latitude = latitude,
    Longitude = longitude,
    Depth_km = depth,
    Magnitude = mag,
    EventID = id
  ) %>%
  mutate(
    Time = as_datetime(Time), # make sure the time col works, I want to use it later
    Source = "USGS" # adding this so it's clear what's usgs and what's texnet
  )

usgs_final <- usgs_clean %>%
  mutate(
    Date = as.Date(Time) # also fix the time column
  ) %>%
  filter(Date <= as.Date("2025-12-31")) %>%
  select(EventID, Date, Latitude, Longitude, Depth_km, Magnitude, Source)

library(sf)
library(tigris)
library(dplyr)

# just have to clip the usgs data from a big square to the actual texas borders

# tx state boundary
# get the texas polygon
tx_boundary <- states(cb = TRUE, class = "sf") %>%
  filter(STUSPS == "TX") %>%
  st_transform(4326)

# convert earthquakes to sf
quakes_sf <- st_as_sf(
  usgs_final, 
  coords = c("Longitude", "Latitude"), 
  crs = 4326,
  remove = FALSE 
)

tx_quakes_sf <- st_filter(quakes_sf, tx_boundary)

usgs_final <- st_drop_geometry(tx_quakes_sf)

fwrite(usgs_final, "Data/usgs_final.csv")
usgs_final <- fread('Data/usgs_final.csv')
```

## Binding the USGS data (2010-2017) to the TexNet Data (2017--)
Seems TexNet has better monitoring and coverage, but it didn't exist until 2017. I still wanted to see seismicity trends as fracking became more prevalent, which is why I use both.
```{r eval= FALSE, echo= TRUE}
library(data.table)
library(dplyr)
library(lubridate)

texnet <- fread('Data/texnet_events.csv')
#cut usgs once texnet starts (texnet has better detection)
usgs_final <- usgs_final %>%
  filter(Date <= as.Date("2016-12-31")) %>%
  select(EventID, Date, Latitude, Longitude, Depth_km, Magnitude, Source)

#  prep texnet data
texnet_final <- texnet %>%
  mutate(
    Date = as.Date(`Origin Date`),
    Source = "TexNet",
    Magnitude = coalesce(`Local Magnitude`, `Moment Magnitude`) # merge the two magnitude columns
  ) %>%
  filter(Date >= as.Date("2017-01-01")) %>%
  select(
    EventID,
    Date,
    Latitude = `Latitude (WGS84)`,
    Longitude = `Longitude (WGS84)`,
    Depth_km = `Depth of Hypocenter (Km. Rel to Ground Surface)`, #  ground surface depth to match USGS standards
    Magnitude,
    Source
  )
fwrite(texnet_final, "Data/texnet_final.csv")
# bind usgs and texnet

master_earthquakes <- bind_rows(usgs_final, texnet_final) %>%
  filter(!is.na(Latitude) & !is.na(Longitude) & !is.na(Magnitude)) %>%
  arrange(Date) 

fwrite(master_earthquakes, "Data/master_earthquakes.csv")

summary(master_earthquakes)
```

```{r eval= FALSE, echo= TRUE}

library(sf)
library(tigris)
library(dplyr)

# tx state boundary
# get the texas polygon
tx_boundary <- states(cb = TRUE, class = "sf") %>%
  filter(STUSPS == "TX") %>%
  st_transform(4326)

# convert earthquakes to sf
quakes_sf <- st_as_sf(
  master_earthquakes, 
  coords = c("Longitude", "Latitude"), 
  crs = 4326,
  remove = FALSE 
)

tx_quakes_sf <- st_filter(quakes_sf, tx_boundary)

master_earthquakes_tx <- st_drop_geometry(tx_quakes_sf)

fwrite(master_earthquakes_tx, "Data/master_earthquakes_tx.csv")
```
# Analysis

## Checking USGS/TexNet Coverage and Examining Earthquake Frequency Trends

```{r eval= TRUE, echo= TRUE, warning= FALSE, message= FALSE}
library(dplyr)
library(ggplot2)
library(data.table)
usgs_final <- fread("data/usgs_final.csv")
texnet_final<- fread("data/texnet_final.csv")

usgs_counts <- usgs_final %>%
  count(Date) %>%
  mutate(source = "USGS")

texnet_counts <- texnet_final %>%
  count(Date) %>%
  mutate(source = "TexNet")

combined_counts <- bind_rows(usgs_counts, texnet_counts)

ggplot(combined_counts, aes(x = Date, y = n, color = source)) +
  # raw daily counts
  geom_line(linewidth = 0.4, alpha = 0.5) +
  # smoothed trend
  geom_smooth(aes(linetype = source),
              method = "loess",
              se = FALSE,
              linewidth = 1.1) +
  # TexNet deployment marker
  geom_vline(xintercept = as.Date("2017-01-01"),
             linetype = "dashed",
             color = "grey40",
             linewidth = 0.7) +
  annotate("text",
           x = as.Date("2017-01-01"),
           y = 50,
           label = "TexNet deployment",
           angle = 90,
           vjust = -.5,
           size = 3.5,
           color = "grey30") +
  
  scale_color_manual(values = c(
    "USGS" = "grey60",
    "TexNet" = "steelblue4")) +
 
   scale_linetype_manual(values = c(
    "USGS" = "solid",
    "TexNet" = "solid" )) +
 
   labs(
    title = "Earthquake Detection in Texas",
    subtitle = "Daily earthquake counts from USGS and TexNet",
    x = "Year",
    y = "Number of Earthquakes",
    color = "Catalog",
    linetype = "Catalog")+
  
  theme_classic() +
  theme(legend.position = "top")+
  ylim(0,100)
```


```{r message= FALSE, echo= FALSE}
#| results: hide
library(sf)
library(tigris)
library(leaflet)
library(dplyr)
library(data.table)

master_frac_cleanest <- fread("data/master_frac_cleanest.csv")
master_earthquakes <- fread("data/master_earthquakes.csv")
#add texas counties
tx_counties <- counties(state = 'TX', cb = TRUE, class = 'sf')
tx_counties <- st_transform(tx_counties, 4326)
```
## Hexbin density plots
```{r}
eq_hex <- ggplot() +
  geom_hex(data = master_earthquakes, aes(x = Longitude, y = Latitude), bins = 40) +
  geom_sf(data = tx_counties, fill = NA, color = "grey40", linewidth = 0.3) +
  
  scale_fill_viridis_c(name = "Earthquake Count") +
  theme_classic()
```

```{r}
frac_hex <- ggplot() +
  geom_hex(data = master_frac_cleanest, aes(x = Longitude, y = Latitude), bins = 40) +
  geom_sf(data = tx_counties, fill = NA, color = "grey40", linewidth = 0.3) +
  scale_fill_viridis_c(name = "Well Count") +
  theme_classic()
```

```{r}
#| echo: false
#| layout-ncol: 2
#| fig-cap: 
#|   - "Density of Fracking Operations"
#|   - "Density of Seismic Events"
frac_hex
eq_hex
```


```{r echo= FALSE, message= FALSE, warning= FALSE}
library(sf)
library(tigris)
library(leaflet)
library(dplyr)
library(data.table)
master_frac_cleanest <- fread("data/master_frac_cleanest.csv")
#wells map
wells_map <- leaflet(options = leafletOptions(preferCanvas = TRUE)) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(
    data = tx_counties,
    fillColor = "lightgrey",
    color = "black",
    weight = 1,
    fillOpacity = 0.2
  ) %>%
  addCircleMarkers(
    data = master_frac_cleanest,
    lng = ~Longitude,
    lat = ~Latitude,
    radius = 3,
    color = "steelblue",
    stroke = FALSE,
    fillOpacity = 0.6,
  )

#wells_map
```

```{r echo= FALSE, message= FALSE, warning= FALSE}
# earthquakes map

master_earthquakes <- fread("data/master_earthquakes.csv")

quakes_map <- leaflet(options = leafletOptions(preferCanvas = TRUE)) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(
    data = tx_counties,
    fillColor = "lightgrey",
    color = "black",
    weight = 1,
    fillOpacity = 0.2
  ) %>%
  addCircleMarkers(
    data = master_earthquakes,
    lng = ~Longitude,
    lat = ~Latitude,
    radius = 3,
    color = "darkred", 
    stroke = FALSE,
    fillOpacity = 0.6,
  )

#quakes_map
```

```{r}
#| echo: false
#| eval: false
library(leafsync)
sync(wells_map, quakes_map)
```

```{r}
col_mag <- colorNumeric('Reds', master_earthquakes$Magnitude)
col_vol <- colorNumeric('Blues', master_frac_cleanest$Total_Fluid_Volume)

wq <- leaflet(options = leafletOptions(preferCanvas = TRUE)) %>% 
  addProviderTiles(providers$CartoDB.DarkMatter) %>%
  addPolygons(
    data = tx_counties,
    fillColor = 'transparent', 
    color = '#444444',        
    weight = 1
  ) %>%

  # wells are plotted first
  addCircleMarkers(
    data = master_frac_cleanest,
    lng = ~Longitude,
    lat = ~Latitude,
    radius = ~(sqrt(Total_Fluid_Volume) / 500) + 1, # this needed special handling because of the range
    color = ~col_vol(Total_Fluid_Volume),   
    stroke = FALSE,
    fillOpacity = 0.15, 
    group = "Fracking Wells"
  ) %>%
  
  # earthquakes sit on top
  addCircleMarkers(
    data = master_earthquakes,
    lng = ~Longitude,
    lat = ~Latitude,
    radius = ~Magnitude *2.5,
    color = ~col_mag(Magnitude),   
    stroke = FALSE,
    fillOpacity = 0.5,  
    group = "Earthquakes"
  ) %>%
  
  addLayersControl(
    overlayGroups = c("Fracking Wells", "Earthquakes"),
    options = layersControlOptions(collapsed = FALSE)
  )

wq
```

## Fracking Volume and Earthquake Frequency
```{r}
frac_month <- master_frac_cleanest %>%
  mutate(month = as.Date(cut(JobEndDate, "month"))) %>%
  group_by(month) %>%
  summarise(total_volume = sum(Total_Fluid_Volume, na.rm=TRUE))

quake_month <- master_earthquakes %>%
  mutate(month = as.Date(cut(Date, "month"))) %>%
  count(month)

combined <- left_join(frac_month, quake_month)
```

```{r}
ggplot(combined, aes(x = month)) +
  geom_line(aes(y = n), color = "red") +
  geom_line(aes(y = total_volume/1e7), color = "steelblue") +
  labs(
    title = "Fracturing Activity vs Earthquake Frequency",
    x = "Year",
    y = "Scaled Values"
  ) +
  theme_classic()
```

# cumulative distribution
```{r}
master_earthquakes %>%
  arrange(Date) %>%
  mutate(cum_eq = row_number()) %>%
  ggplot(aes(Date, cum_eq)) +
  geom_line(color="firebrick") +
  theme_classic()
```


```{r echo= FALSE, eval= FALSE}
ggplot(master_frac_cleanest,
       aes(TVD, Total_Fluid_Volume)) +
  geom_point(alpha = .2) +
  geom_smooth(method = "loess") +
  scale_y_log10() +
  theme_classic()
```


# References