---
title: "Financial Structure, Patent Protection, and the Quality–Quantity Tradeoff in Innovation"
subtitle: "Replication Appendix"
author: Aaron DiLorenzo
date: today
format:
  html:
    linkcolor: '#74a7d8'
    theme: default
    page-layout: article
    toc: true
    toc-depth: 1
    toc-location: right
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    code-link: true
    code-overflow: scroll
    code-line-numbers: true
    code-copy: true
    include-in-header:
        text: |
          <link rel="stylesheet" type="text/css" href="https://cdn.jsdelivr.net/gh/bitmaks/cm-web-fonts@latest/fonts.css">
          <style>
            body, p, h1, h2, h3, h4, h5, h6, li, td, th {
              font-family: "Computer Modern Serif", serif !important;
            }
          </style>
execute:
  enabled: true
  python: false
  
  echo: true      
  eval: true     
  warning: false
  message: true
  cache: true
  freeze: auto
number-sections: true
self-contained: true  
editor: 
  markdown: 
    wrap: 72
---
---

### Resources & Data Access

* **[Download/View the Full Thesis (PDF)](documents/analysis.pdf)** 
* **[Download the Replication Source Code (.qmd)](https://github.com/diloaaron/diloaaron.github.io/blob/main/thesis.qmd)**
* **[Access the Raw and Final Datasets](https://github.com/diloaaron/diloaaron.github.io/tree/main/data)**
* **[Request OECD Patent Quality Data](https://forms.office.com/pages/responsepage.aspx?id=1MdBrGEfDUaw9PySWitHHKuxmuqpz_9KusL7-G1D6wFUOEU0OVBYVk5QTzROVlBTSUtBUUREWVhHTiQlQCN0PWcu&route=shorturl)** *(Required for the initial data construction phase)*

---

# Patent Quality Indicators Construction

The dependent variables for patent quality are derived from the OECD Patent Quality Indicators database (2025 edition). While the database provides a wide array of metrics (forward citations, family size, and grant lag), this thesis focuses on two specific measures:

Composite Quality Index (`quality_index_4`): An unweighted average of four normalized components (forward citations, patent family size, radicalness, and number of technical fields).

Breakthrough Patents (`breakthrough`): A binary indicator identifying patents in the top 1% of the citation distribution within their respective technology-year cohort.

Because the OECD quality indicators are provided at the patent-application level without geographical identifiers, a multi-stage merging procedure is required to construct the country-year panel used in the analysis. The Quality Indicators file is joined with the OECD HAN (Harmonised Applicant Names) database using the unique application ID (`appln_id`). To assign each patent to a country, the HAN database is linked to the Harmonised Names (HARM_NAMES) database, allowing me to map each application to the `Person_ctry_code` (ISO country code) of the patent applicant. Patents are assigned to years based on their filing date. Finally, the patent-level data is aggregated to the country-year level. The `quality_index_4` is averaged across all patents for a given country-year to create a mean quality score, while the breakthrough indicator is summed to calculate the total count of high-impact patents.

The following Python script (using the `polars` library for high-performance data manipulation) executes this pipeline:

```{python, eval=FALSE}
"""
OECD Patent Quality Indicators
normalization, winsorization, and aggregation procedures
"""

import polars as pl
import os

OUTPUT_DIR = ""
os.makedirs(OUTPUT_DIR, exist_ok=True)
# for reproducing, change the file paths to correctly the downloaded data, 
# they're too large for me to make this self running
INDICATORS_FILE = "Data/202502_OECD_PATENT_QUALITY_EPO_INDIC.txt"
COHORTS_FILE = "Data/202502_OECD_PATENT_QUALITY_EPO_INDIC_COHORT.txt"
HAN_PATENTS_FILE = "Data/202502_HAN_PATENTS.txt"
HARM_NAMES_FILE = "Data/202502_HARM_NAMES.txt"

FINAL_PATENT_FILE = "OECD_patent_quality_final.parquet"
FINAL_COUNTRY_FILE = "OECD_country_year_final.csv"

WINSOR_VARS = ["family_size", "grant_lag", "bwd_cits", "npl_cits", "claims", "fwd_cits5", "fwd_cits7"]
COMPOSITE_VARS = ["generality", "originality", "radicalness", "quality_index_4", "quality_index_6"]

# load the data

def load_data():

    cohorts = pl.read_csv(COHORTS_FILE, separator="|", infer_schema_length=0)
    cohorts = cohorts.rename({c: c.strip() for c in cohorts.columns})

    indicators = pl.read_csv(INDICATORS_FILE, separator="|", infer_schema_length=10000)
    indicators = indicators.rename({c: c.strip() for c in indicators.columns})

    han = pl.read_csv(HAN_PATENTS_FILE, separator="|")
    han = han.rename({c: c.strip() for c in han.columns})
    han = han.rename({"Appln_id": "appln_id"})

    harm = pl.read_csv(HARM_NAMES_FILE, separator="|")
    harm = harm.rename({c: c.strip() for c in harm.columns})
    harm = harm.select(["HARM_ID", "Person_ctry_code"])

    han = han.join(harm, on="HARM_ID", how="left")

    return cohorts, indicators, han

# main function

def main():

    cohorts, indicators, han = load_data()

    merged = indicators.join(han, on="appln_id", how="left")

    # ensure breakthrough is binary
    if "breakthrough" in merged.columns:
        merged = merged.with_columns(
            pl.col("breakthrough").fill_null(0).cast(pl.Int8)
        )

    # aggregation to country/year

    agg_exprs = []

    # normalized indicators
    norm_cols = [c for c in merged.columns if c.endswith("_norm")]
    for c in norm_cols:
        agg_exprs.append(pl.mean(c).alias(f"avg_{c}"))

    # mean for composite indicator
    for c in COMPOSITE_VARS:
        if c in merged.columns:
            agg_exprs.append(pl.mean(c).alias(f"avg_{c}"))

    # breakthrough counts
    agg_exprs.extend([
        pl.sum("breakthrough").alias("breakthrough_count"),
        pl.count("appln_id").alias("patent_count"),
        (pl.sum("breakthrough") / pl.count("appln_id")).alias("pct_breakthrough")
    ])

    country_agg = (
        merged
        .filter(pl.col("Person_ctry_code").is_not_null())
        .group_by(["Person_ctry_code", "filing"])
        .agg(agg_exprs)
        .rename({"Person_ctry_code": "iso3c", "filing": "year"})
        .sort(["iso3c", "year"])
    )

    country_agg.write_csv(
        os.path.join(OUTPUT_DIR, FINAL_COUNTRY_FILE)
    )

    merged.write_parquet(
        os.path.join(OUTPUT_DIR, FINAL_PATENT_FILE)
    )

if __name__ == "__main__":
    main()
```

# Data Construction and Variable Assembly

In this section, I assemble the final panel dataset by merging the OECD
Patent Quality Indicators with macroeconomic and financial development
data.

### Dependent Variables

I construct two primary dependent variables: 
1. **Patent Quality (`quality_index_100`)**: The raw OECD composite quality index is scaled
to a 0-100 range for ease of interpretation. 

2. **Patent Quantity (`ln_patents_pc`)**: patenting
intensity is measured as patents per capita (scaled per million
residents). This is then $\log(x+1)$ transformed to account for
right-skewness and potential zero values.

### Independent Variables & Mean-Centering

To accurately capture the interaction between financial structure and
intellectual property rights, I apply a $\log(x+1)$ transformation to
our financial development variables (Bank Credit, Market Capitalization,
and Venture Capital as a % of GDP) due to their right-skewness.

**Both** the Park Index (IPR) and the logged Financial Development
variables are mean-centered *prior* to interaction. This ensures that
the main effect coefficients in the regressions represent the
marginal effect evaluated at the sample mean, significantly improving
the interpretability of the final models.

```{r data-assembly, eval=TRUE, echo=TRUE, message= FALSE, warning= FALSE}
library(data.table)
library(dplyr)

options(dplyr.summarise.inform = FALSE)

# 1. load individual datasets
pqi      <- fread("data/OECD_country_year_final.csv")
fd       <- fread("data/FD_Final.csv")
vc       <- fread("data/VC_final.csv")
counts   <- fread("data/counts_res_final.csv")
controls <- fread("data/Controls_final.csv")
park     <- fread("data/Park_final.csv")

# 2. clean up and standardize missing values
clean_missing <- function(df) {
  df[df == ".."] <- NA
  df[df == "NA"] <- NA
  df[df == ""]   <- NA
  return(df)
}

pqi      <- clean_missing(pqi)
fd       <- clean_missing(fd)
vc       <- clean_missing(vc)
counts   <- clean_missing(counts)
controls <- clean_missing(controls)
park     <- clean_missing(park)

# Remove old 'researchers' column to prevent any errors
counts <- counts %>% select(-researchers_per_million)

# 3. merge individual datasets into a master panel
df <- pqi %>%
  left_join(park %>% select(iso3c, year, ipp_index_IP_transformed), by = c("iso3c", "year")) %>%
  left_join(fd,       by = c("iso3c", "year")) %>%
  left_join(vc,       by = c("iso3c", "year")) %>%
  left_join(counts,   by = c("iso3c", "year")) %>%
  left_join(controls, by = c("iso3c", "year"))

# 4. ensure correct numeric typing
numeric_vars <- c(
  "population_total", "gdp_pc_const2015usd", "bank_credit_gdp", 
  "market_cap_gdp", "VC_investment", "trade_percent_gdp", 
  "rd_expenditure_percent_gdp", "tertiary_enrollment_percent", 
  "patent_applications_residents", "breakthrough_count", 
  "patent_count", "ipp_index_IP_transformed"
)

for (v in numeric_vars) {
  if (v %in% names(df)) {
    df[[v]] <- as.numeric(df[[v]])
  }
}

# 5. Variable Construction and Transformations
df <- df %>%
  mutate(
    # dependent variable: Patents per capita (scaled per million residents)
    patents_per_capita = (patent_applications_residents / population_total) * 1000000,
    patents_per_capita = ifelse(
      is.infinite(patents_per_capita) | patents_per_capita < 0,
      NA,
      patents_per_capita
    ),
    
    # dependent variable: Scaled Patent Quality
    quality_index_100 = avg_quality_index_4 * 100,
    
    # rescale VC to match the percentage format of other financial variables
    VC_investment_pct = VC_investment * 100,
    
    # log transformations (log(x + 1))
    ln_patents_pc = log(patents_per_capita + 1),
    ln_gdp_pc     = log(gdp_pc_const2015usd + 1),
    ln_bank       = log(bank_credit_gdp + 1),
    ln_market     = log(market_cap_gdp + 1),
    ln_vc         = log(VC_investment_pct + 1),
    ln_trade      = log(trade_percent_gdp + 1),
    ln_rd         = log(rd_expenditure_percent_gdp + 1),
    ln_tertiary   = log(tertiary_enrollment_percent + 1),
    
    # mean-centering prior to interaction
    ipr_c       = ipp_index_IP_transformed - mean(ipp_index_IP_transformed, na.rm = TRUE),
    ln_bank_c   = ln_bank - mean(ln_bank, na.rm = TRUE),
    ln_market_c = ln_market - mean(ln_market, na.rm = TRUE),
    ln_vc_c     = ln_vc - mean(ln_vc, na.rm = TRUE),
    
    # interaction terms
    ln_bank_x_ipr   = ln_bank_c * ipr_c,
    ln_market_x_ipr = ln_market_c * ipr_c,
    ln_vc_x_ipr     = ln_vc_c * ipr_c
  ) %>%
  # drop unneeded intermediate columns
  select(
    -avg_generality, -avg_originality, -avg_radicalness, 
    -`Country Name.x`, -FS.AST.PRVT.GD.ZS, -CM.MKT.LCAP.GD.ZS, 
    -`Country Name.y`
  )
```

# Baseline Two-Way Fixed Effects (TWFE) Models

In this section, I estimate two-way fixed effects (TWFE) panel
regressions to establish baseline relationships between financial
structure, intellectual property rights, and my two distinct innovation
outcomes: Patent Quality and Patent Quantity.

### Model Specification and Methodological Choices

For each of the three financial development channels (Market
Capitalization, Bank Credit, and Venture Capital), I estimate models
sequentially introducing macroeconomic controls.

The baseline equation takes the form:
$$Y_{it} = \beta_1 FD_{it}^c + \beta_2 IPR_{it}^c + \beta_3 (FD_{it}^c \times IPR_{it}^c) + \mathbf{X'}_{it}\beta + \alpha_i + \lambda_t + \epsilon_{it}$$

**Features of this approach:** 1. **Centered Variables (**$^c$): As
established in the data preparation phase, both the financial
development variable ($FD$) and the Park Index ($IPR$) are
mean-centered. This ensures that $\beta_1$ and $\beta_2$ are interpreted
as the marginal effects of finance and IPR exactly at the sample mean of
the other variable, rather than at an unrealistic zero value. 2. **Fixed
Effects (**$\alpha_i, \delta_t$): I include country fixed effects to
absorb time-invariant institutional or cultural characteristics, and
year fixed effects to control for global macro shocks and
aggregate patenting trends. 3. **Clustered Standard Errors**: To account
for serial correlation within countries over time, all standard errors
are clustered at the country level using `vcovHC`.

### Estimation and HTML Tables

The following custom function estimates the incremental models and
outputs clean HTML tables using `stargazer`. Note that the dataset `df`
is carried over directly from the previous chunk's environment.

```{r twfe-models, results='asis', echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}
library(plm)
library(lmtest)
library(sandwich)
library(stargazer)

# function to calculate country-clustered standard errors
cluster_se <- function(model) {
  sqrt(diag(vcovHC(model, type = "HC1", cluster = "group")))
}

# master function to run TWFE models and output HTML tables
run_twfe_incremental <- function(depvar, fd, fd_x_ipr, title) {

  # filter missing values specific to the target variables to maximize sample size
  base <- df %>%
    filter(
      !is.na(.data[[depvar]]),
      !is.na(.data[[fd]]),
      !is.na(ipr_c)
    )

  # convert to panel dataframe
  panel <- pdata.frame(base, index = c("iso3c", "year"))
  
  # define sequential formulas
  f1 <- as.formula(paste(depvar, "~", fd, "+ ipr_c +", fd_x_ipr))
  f2 <- update(f1, . ~ . + ln_gdp_pc)
  f3 <- update(f2, . ~ . + ln_trade)
  f4 <- update(f3, . ~ . + ln_rd)
  f5 <- update(f4, . ~ . + ln_tertiary)

  # estimate within-country, two-way models
  m1 <- plm(f1, data = panel, model = "within", effect = "twoways")
  m2 <- plm(f2, data = panel, model = "within", effect = "twoways")
  m3 <- plm(f3, data = panel, model = "within", effect = "twoways")
  m4 <- plm(f4, data = panel, model = "within", effect = "twoways")
  m5 <- plm(f5, data = panel, model = "within", effect = "twoways")
  
  # format labels
  fin_label <- switch(fd,
    "ln_market_c" = "Log Market Cap (Centered)",
    "ln_bank_c"   = "Log Bank Credit (Centered)",
    "ln_vc_c"     = "Log Venture Capital (Centered)"
  )
  
  # output formatted HTML table for this document
  html_table <- capture.output(
    stargazer(
      m1, m2, m3, m4, m5,
      type = "html",
      title = title,
      dep.var.caption = ifelse(
        depvar == "quality_index_100",
        "Dependent Variable: Patent Quality (0–100)",
        "Dependent Variable: Log Patents per Capita"
      ),
      covariate.labels = c(
        fin_label,
        "IPR Index (Centered)",
        paste(fin_label, "x IPR"),
        "Log GDP per Capita",
        "Log Trade Openness",
        "Log R&D Expenditure",
        "Log Tertiary Enrollment"
      ),
      se = list(
        cluster_se(m1), cluster_se(m2), cluster_se(m3), cluster_se(m4), cluster_se(m5)
      ),
      omit.stat = c("ser", "f"),
      notes = c(
        "Country and year fixed effects included in all models.",
        "Standard errors clustered by country."
      ),
      notes.append = TRUE,
      no.space = TRUE,
      header = FALSE 
    )
  )

  # print HTML table to the Quarto render
  cat(paste(html_table, collapse = "\n"))
  cat("<br><br>") # Add spacing between tables

  invisible(list(m1, m2, m3, m4, m5))
}

# patent quality models

market_quality <- run_twfe_incremental(
  depvar   = "quality_index_100",
  fd       = "ln_market_c",       
  fd_x_ipr = "ln_market_x_ipr",   
  title    = "Market Finance and Average Patent Quality"
)

bank_quality <- run_twfe_incremental(
  depvar   = "quality_index_100",
  fd       = "ln_bank_c",         
  fd_x_ipr = "ln_bank_x_ipr",     
  title    = "Bank Finance and Average Patent Quality"
)

vc_quality <- run_twfe_incremental(
  depvar   = "quality_index_100",
  fd       = "ln_vc_c",           
  fd_x_ipr = "ln_vc_x_ipr",       
  title    = "Venture Capital and Average Patent Quality"
)

# patent quantity models

market_quantity <- run_twfe_incremental(
  depvar   = "ln_patents_pc",
  fd       = "ln_market_c",       
  fd_x_ipr = "ln_market_x_ipr",
  title    = "Market Finance and Patent Quantity (per capita)"
)

bank_quantity <- run_twfe_incremental(
  depvar   = "ln_patents_pc",
  fd       = "ln_bank_c",         
  fd_x_ipr = "ln_bank_x_ipr",
  title    = "Bank Finance and Patent Quantity (per capita)"
)

vc_quantity <- run_twfe_incremental(
  depvar   = "ln_patents_pc",
  fd       = "ln_vc_c",           
  fd_x_ipr = "ln_vc_x_ipr",
  title    = "Venture Capital and Patent Quantity (per capita)"
)
```

# Zero-Inflated Negative Binomial (ZINB) Models

In this supplementary analysis, I shift the focus from average patent quality and volume to the extreme right tail of the innovation distribution: **breakthrough patents**. 

### Methodological Justification

Breakthrough innovations are inherently rare, meaning the `breakthrough_count` variable is strictly non-negative count data characterized by two econometric challenges:

1. **Excess Zeros**: Most country-year observations record zero breakthrough patents.

2. **Overdispersion**: The variance of breakthrough counts significantly exceeds the mean, violating the assumption of standard Poisson regression.

To address this, I estimate Zero-Inflated Negative Binomial (ZINB) models. The ZINB framework simultaneously estimates two equations:

* **The Selection (Zero-Inflation) Model**: A logistic regression predicting the probability that a country is in a "certain zero" state.

* **The Count Model**: A negative binomial regression estimating the actual number of breakthrough patents produced, conditional on the country clearing the zero-hurdle, while explicitly modeling the overdispersion.

### Model Specification

Just as in the baseline models, I utilize mean-centered financial development ($FD_{it}^c$) and institutional quality ($IPR_{it}^c$) variables. The models below iterate through the financial channels (Market Capitalization, Bank Credit, and Venture Capital), sequentially adding macroeconomic controls. 

*(Note: Because ZINB models generate coefficients for both the count and zero-inflation equations simultaneously, the raw model outputs are provided below. Formatted, consolidated tables are available in the main text of the thesis).*

```{r zinb-models, results='asis', echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}
library(pscl)
library(dplyr)
library(knitr)

# create the offset variable (internal patents from the OECD data, distinct from constructed patents/capita)

df$ln_internal_patents <- log(df$patent_count)

# function to run ZINB and output clean HTML tables
run_zinb_incremental <- function(finance_var, title_name) {

  # match exactly with the variable names 
  fd_c  <- paste0(finance_var, "_c")
  fd_x  <- paste0(finance_var, "_x_ipr") 

  controls <- c("ln_gdp_pc", "ln_trade", "ln_rd", "ln_tertiary")
  models <- list()

  # 1. estimate Models
  for (i in 1:4) {
    intensity_controls <- paste(controls[1:i], collapse = " + ")

    formula_str <- paste0(
      "breakthrough_count ~ ",
      fd_c, " + ipr_c + ", fd_x,
      " + ", intensity_controls,
      " + offset(ln_internal_patents) | ln_gdp_pc"
    )

    models[[i]] <- zeroinfl(
      as.formula(formula_str),
      data = df,
      dist = "negbin"
    )
  }

  # 2. output HTML Tables
  cat("##", title_name, "\n\n")

  for (j in 1:4) {
    cat("### Model", j, "\n\n")
    model_sum <- summary(models[[j]])

    # calculate Log Theta explicitly for the HTML table
    m <- models[[j]]
    if (!is.null(m$theta) && !is.null(m$SE.theta)) {
      lt_est <- log(m$theta)
      lt_se  <- m$SE.theta / m$theta
    } else {
      lt_est <- NA
      lt_se  <- NA
    }

    # Count model
    count_coefs <- model_sum$coefficients$count
    if (!is.na(lt_est)) {
      count_coefs <- rbind(count_coefs, "Log Theta" = c(lt_est, lt_se, NA, NA))
    }

    count_table <- data.frame(
      Variable = rownames(count_coefs),
      Coefficient = round(count_coefs[,1], 3),
      Std_Error = round(count_coefs[,2], 3)
    )

    cat("#### Intensity (Negative Binomial Count Equation)\n\n")
    print(knitr::kable(count_table, row.names = FALSE))
    cat("\n\n")

    # selection model
    zero_coefs <- model_sum$coefficients$zero
    zero_table <- data.frame(
      Variable = rownames(zero_coefs),
      Coefficient = round(zero_coefs[,1], 3),
      Std_Error = round(zero_coefs[,2], 3)
    )

    cat("#### Selection (Logit Structural Zero Equation)\n\n")
    print(knitr::kable(zero_table, row.names = FALSE))
    cat("\n\n")

    cat("**Observations:**", models[[j]]$n, "\n\n")
    cat("---\n\n") 
  }

  invisible(models)
}

# execute the models
zinb_market <- run_zinb_incremental("ln_market", "ZINB: Market Finance and Breakthrough Innovation")

zinb_bank <- run_zinb_incremental("ln_bank", "ZINB: Bank Finance and Breakthrough Innovation")

zinb_vc <- run_zinb_incremental("ln_vc", "ZINB: Venture Capital and Breakthrough Innovation")
```

# Data Diagnostics and Exploratory Analysis

## Breakthrough Innovation Measure

```{r eval= TRUE, echo= TRUE, message= FALSE, warning= FALSE}
#| column: margin
library(ggplot2)
library(dplyr)

# breakthrough count diagnostics (justifies model choice)



cat("Total observations:", nrow(df), "\n")
cat("Zero breakthrough observations:", sum(df$breakthrough_count == 0, na.rm = TRUE), "\n")
cat("Percent zeros:", round(mean(df$breakthrough_count == 0, na.rm = TRUE) * 100, 2), "%\n")

cat("Mean breakthrough count:", mean(df$breakthrough_count, na.rm = TRUE), "\n")
cat("Variance breakthrough count:", var(df$breakthrough_count, na.rm = TRUE), "\n")

breakthrough_dstb <- ggplot(df, aes(x = breakthrough_count)) +
  geom_histogram(binwidth = 1,
                 fill = "#2C3E50",
                 color = "white",
                 linewidth = 0.2) +
  coord_cartesian(xlim = c(0, 30)) +
  labs(
    title = "Distribution of Breakthrough Patent Counts",
    x = "Breakthrough Count (Top 1% Patents)",
    y = "Frequency"
  ) +
  theme_classic(base_family = "serif", base_size = 12) +
  theme(
    plot.title = element_text(face = "bold"),
    axis.title = element_text(face = "plain")
  )
breakthrough_dstb
# ggsave("figures/breakthrough_distribution.pdf",
#        plot = breakthrough_dstb,
#        width = 7,
#        height = 5)
```

## Distribution of Patent Quality

```{r eval= TRUE, echo= TRUE, message= FALSE, warning= FALSE}
#| column: margin

quality_density <- ggplot(df, aes(x = quality_index_100)) +
  geom_density(fill = "#2C3E50", alpha = 0.4, color = "#2C3E50", linewidth = 0.6) +
  labs(
    title = "Distribution of Patent Quality Index",
    x = "Patent Quality (0-100)",
    y = "Density"
  ) +
  theme_classic(base_family = "serif", base_size = 12) +
  theme(
    plot.title = element_text(face = "bold"),
    axis.title = element_text(face = "plain")
  )
quality_density
# ggsave("figures/quality_density.pdf",
#        plot = quality_density,
#        width = 7,
#        height = 5)
```

## Distribution of Financial Development Variables

```{r fig.width= 15, fig.height= 5, echo= TRUE, eval= TRUE, warning= FALSE, message= FALSE}
#| column: margin
library(dplyr)
library(tidyr)
library(ggplot2)

vars <- c("ln_market_c", "ln_bank_c", "ln_vc_c")

df_long <- df %>%
  select(all_of(vars)) %>%
  pivot_longer(everything()) %>%
  mutate(name = dplyr::recode(name,
                       ln_market_c = "Market Finance",
                       ln_bank_c   = "Bank Finance",
                       ln_vc_c     = "VC Finance"))

finance_density <- ggplot(df_long, aes(x = value)) +
  geom_density(fill = "#2C3E50",
               alpha = 0.4,
               color = "#2C3E50",
               linewidth = 0.6) +
  facet_wrap(~name, scales = "free") +
  labs(
    title = "Distribution of Financial Development Measures",
    x = "Log of Financial Development (% of GDP)",
    y = "Density"
  ) +
  theme_classic(base_family = "serif", base_size = 12) +
  theme(
    plot.title = element_text(face = "bold"),
    strip.text = element_text(face = "bold")
  )
finance_density
# ggsave("figures/finance_distribution.pdf",
#        plot = finance_density,
#        width = 8,
#        height = 5)
```

## Correlation Matrix

```{r}
#| fig-width: 8
#| fig-height: 8
#| column: margin
library(dplyr)
library(corrplot)

corr_vars <- df %>%
  select(quality_index_100, ln_patents_pc,
         ln_market_c, ln_bank_c, ln_vc_c,
         ipr_c, ln_gdp_pc, ln_trade,
         ln_rd, ln_tertiary) %>%
  na.omit() %>%
  rename(
    "Patent Quality" = quality_index_100,
    "Log Patents" = ln_patents_pc,
    "Market Finance" = ln_market_c,
    "Bank Finance" = ln_bank_c,
    "VC Finance" = ln_vc_c,
    "IPR Index" = ipr_c,
    "Log GDP per Capita" = ln_gdp_pc,
    "Log Trade" = ln_trade,
    "Log R&D" = ln_rd,
    "Log Tertiary Education" = ln_tertiary
  )

corr_matrix <- cor(corr_vars)

par(family = "serif")

corrplot(
  corr_matrix,
  method = "color",
  type = "upper",
  diag = FALSE,
  col = colorRampPalette(c("#2C3E50", "white", "#8B2E2E"))(200),
  tl.col = "black",
  tl.cex = 0.9,
  number.cex = 0.75,
  addCoef.col = "black"
)

# dev.copy(pdf, "figures/correlation_matrix.pdf", width = 8, height = 8)
# dev.off()
```

## Interaction Effects (Marginal Effects Plot)

```{r echo= TRUE, eval= TRUE, message= FALSE, warning= FALSE}
#| column: margin
library(plm)
library(marginaleffects)
library(ggplot2)
library(dplyr)

# -------------------------
# Estimate Column 3
# -------------------------
twfe_c3 <- plm(
  quality_index_100 ~ 
    ln_market_c * ipr_c +
    ln_gdp_pc +
    ln_trade,
  data = df,
  index = c("iso3c", "year"),
  model = "within",
  effect = "twoways"
)

# -------------------------
# Estimate Column 4
# -------------------------
twfe_c4 <- plm(
  quality_index_100 ~ 
    ln_market_c * ipr_c +
    ln_gdp_pc +
    ln_trade +
    ln_rd,
  data = df,
  index = c("iso3c", "year"),
  model = "within",
  effect = "twoways"
)

# Sequence over centered ln_market range
ln_seq <- seq(
  min(df$ln_market_c, na.rm = TRUE),
  max(df$ln_market_c, na.rm = TRUE),
  length.out = 200
)

# marginal effects (col 3)
mfx_c3 <- slopes(
  twfe_c3,
  variables = "ipr_c",
  newdata = datagrid(ln_market_c = ln_seq),
  vcov = "HC1"
) %>%
  mutate(model = "Column 3")

#marginal effects (col 4)
mfx_c4 <- slopes(
  twfe_c4,
  variables = "ipr_c",
  newdata = datagrid(ln_market_c = ln_seq),
  vcov = "HC1"
) %>%
  mutate(model = "Column 4")

# Combine
mfx_all <- bind_rows(mfx_c3, mfx_c4)

#turning points
coef_c3 <- coef(twfe_c3)
coef_c4 <- coef(twfe_c4)

tp_c3 <- -coef_c3["ipr_c"] / coef_c3["ln_market_c:ipr_c"]
tp_c4 <- -coef_c4["ipr_c"] / coef_c4["ln_market_c:ipr_c"]

# plot
p <- ggplot(mfx_all, aes(x = ln_market_c, y = estimate)) +
  geom_line(size = 1.1, color = "black") +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high),
              alpha = 0.2, fill = "gray70") +
  geom_hline(yintercept = 0, linetype = "dashed") +
  facet_wrap(~ model) +
  geom_vline(
    data = data.frame(model = c("Column 3","Column 4"),
                      tp = c(tp_c3, tp_c4)),
    aes(xintercept = tp),
    linetype = "dashed"
  ) +
  labs(
    title = "Marginal Effect of IPR on Patent Quality",
    subtitle = 'Column 3: -0.737 | Column 4: -0.088',
    x = "Log Market Capitalization (Centered)",
    y = "Marginal Effect of IPR on Quality"
  ) +
  theme_classic(base_family = "serif") +
  theme(
    strip.text = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 14),
    plot.title = element_text(size = 16, face = "bold")
  )

#ggsave("figures/interaction_quality2.pdf", plot = p, width = 10, height = 5)

p
```

## Additional Figures

```{r eval= TRUE, echo= TRUE, message= FALSE, warning= FALSE}
#| column: margin
# distibution of patent quality
library(tidyverse)
library(data.table)


# create bins
df_plot <- df %>%
  filter(!is.na(ln_market_c), !is.na(ipr_c), !is.na(quality_index_100)) %>%
  mutate(
    market_bin = ntile(ln_market, 3),
    ipr_regime = ifelse(ipr_c <= median(ipr_c, na.rm = TRUE),
                        "Weak IPR", "Strong IPR"),
    group = paste0("Market Q", market_bin, " | ", ipr_regime)
  )

quality_density_bins <- ggplot(df_plot, aes(x = quality_index_100, fill = ipr_regime)) +
  geom_density(alpha = 0.4, color = 'grey30') +
  
  facet_wrap(~ market_bin, nrow = 1,
             labeller = labeller(market_bin = function(x) paste("Market Tercile", x))) +
  
  scale_fill_manual(
    values = c("Weak IPR" = "grey70",
               "Strong IPR" = "#2C3E50")
  ) +
  
  labs(
    title = "Innovation Quality by Market Finance and IPR Regime",
    x = "Patent Quality Index (0-100)",
    y = "Density",
    fill = "IPR Regime"
  ) +
  
  theme_classic(base_family = "serif", base_size = 12) +
  theme(
    plot.title = element_text(face = "bold"),
    legend.position = "top",
    strip.text = element_text(face = "bold")
  )

# ggsave(
#   "figures/quality_by_market_ipr.pdf",
#   plot = quality_density_bins,
#   width = 7,
#   height = 5
# )
# quality and quantity relationship
p_quality_quantity <- ggplot(df, aes(x = ln_patents_pc, y = quality_index_100)) +
  geom_point(alpha = 0.25, color = "grey50") +
  
  geom_smooth(
    method = "loess",
    se = FALSE,
    color = "#2C3E50",
    linewidth = 1.1
  ) +
  
  labs(
    title = "Patent Quantity and Average Patent Quality",
    x = "Log Patent Quantity",
    y = "Average Patent Quality"
  ) +
  
  theme_classic(base_family = "serif", base_size = 12) +
  theme(
    plot.title = element_text(face = "bold")
  )

# ggsave(
#   "figures/quality_vs_quantity.pdf",
#   plot = p_quality_quantity,
#   width = 7,
#   height = 5
# )

quality_density_bins
p_quality_quantity
```

```{r eval= TRUE, echo= TRUE, message= FALSE, warning= FALSE, fig.width= 10, fig.height= 5}
#| column: margin
library(tidyverse)

plot_df <- df %>%
  select(
    iso3c, year,
    quality_index_100,
    ln_patents_pc,
    ipr_c,
    ln_bank_c, ln_market_c, ln_vc_c
  ) %>%
  pivot_longer(
    cols = c(ln_bank_c, ln_market_c, ln_vc_c),
    names_to = "finance_type",
    values_to = "finance_depth"
  ) %>%
  mutate(
    finance_type = case_match(
      finance_type,
      "ln_bank_c"   ~ "Bank Finance",
      "ln_market_c" ~ "Market Finance",
      "ln_vc_c"     ~ "Venture Capital",
      .default = finance_type
    )
  )

# finance x quality

p1 <- ggplot(plot_df, aes(x = finance_depth, y = quality_index_100)) +
  geom_point(alpha = 0.4, color = "grey50") +
  geom_smooth(method = "loess", se = FALSE,
              linewidth = 1.1, color = "#2C3E50") +
  facet_wrap(~ finance_type, scales = "free_x") +
  labs(
    title = "Financial Development and Patent Quality",
    x = "Log Financial Development (% of GDP)",
    y = "Patent Quality Index (0-100)"
  ) +
  theme_classic(base_family = "serif", base_size = 12) +
  theme(
    plot.title = element_text(face = "bold"),
    strip.text = element_text(face = "bold")
  )

# ggsave("figures/finance_quality.pdf", plot = p1, width = 7, height = 5)

# finance x quantity

p2 <- ggplot(plot_df, aes(x = finance_depth, y = ln_patents_pc)) +
  geom_point(alpha = 0.4, color = "grey50") +
  geom_smooth(method = "loess", se = FALSE,
              linewidth = 1.1, color = "#2C3E50") +
  facet_wrap(~ finance_type, scales = "free_x") +
  labs(
    title = "Financial Development and Patent Quantity",
    x = "Log Financial Development (% of GDP)",
    y = "Log Patents per 1,000 Researchers"
  ) +
  theme_classic(base_family = "serif", base_size = 12) +
  theme(
    plot.title = element_text(face = "bold"),
    strip.text = element_text(face = "bold")
  )

# ggsave("figures/finance_quantity.pdf", plot = p2, width = 7, height = 5)

# ipr x quality

p3 <- ggplot(df, aes(x = ipr_c, y = quality_index_100)) +
  geom_point(alpha = 0.4, color = "grey50") +
  geom_smooth(method = "loess", se = FALSE,
              linewidth = 1.1, color = "#2C3E50") +
  labs(
    title = "IPR Strength and Patent Quality",
    x = "IPR Strength (Centered Park Index)",
    y = "Patent Quality Index (0-100)"
  ) +
  theme_classic(base_family = "serif", base_size = 12) +
  theme(
    plot.title = element_text(face = "bold")
  )

# ggsave("figures/ipr_quality.pdf", plot = p3, width = 7, height = 5)

# ipr x quantity

p4 <- ggplot(df, aes(x = ipr_c, y = ln_patents_pc)) +
  geom_point(alpha = 0.4, color = "grey50") +
  geom_smooth(method = "loess", se = FALSE,
              linewidth = 1.1, color = "#2C3E50") +
  labs(
    title = "IPR Strength and Patent Quantity",
    x = "IPR Strength (Centered Park Index)",
    y = "Log Patents per Capita"
  ) +
  theme_classic(base_family = "serif", base_size = 12) +
  theme(
    plot.title = element_text(face = "bold")
  )

# ggsave("figures/ipr_quantity.pdf", plot = p4, width = 7, height = 5)
p1
p2
p3
p4
```

```{r eval= TRUE, echo= TRUE, message= FALSE, warning= FALSE}
#| column: margin
quality_market_df <- df %>%
  filter(
    !is.na(quality_index_100),
    !is.na(ln_market_c),
    !is.na(ipr_c)
  )

quality_market_ipr_df <- quality_market_df %>%
  mutate(
    ipr_regime = ifelse(ipr_c >= 0,
                        "Above-Average IPR",
                        "Below-Average IPR")
  )

p_quality_market_ipr <- ggplot(
  quality_market_ipr_df,
  aes(x = ln_market_c,
      y = quality_index_100,
      color = ipr_regime)
) +
  
  geom_point(alpha = 0.3) +
  
  geom_smooth(method = "lm",
              se = FALSE,
              linewidth = 1.2) +
  
  scale_color_manual(
    values = c("Below-Average IPR" = "grey60",
               "Above-Average IPR" = "#2C3E50")
  ) +
  
  labs(
    title = "Patent Quality, Market Finance, and IPR Regimes",
    x = "Log Market Capitalization (% of GDP)",
    y = "Patent Quality Index (0-100)",
    color = "IPR Regime:"
  ) +
  
  theme_classic(base_family = "serif", base_size = 12) +
  theme(
    plot.title = element_text(face = "bold"),
    legend.position = "top"
  )

# ggsave(
#   "figures/quality_market_ipr.pdf",
#   plot = p_quality_market_ipr,
#   width = 7,
#   height = 5
# )
p_quality_market_ipr
```

# Robustness Checks


## Joint Significance Test

# Robustness Checks

## Joint Significance Tests for Interaction Models

To confirm that these variables meaningfully impact our innovation outcomes, I perform joint significance (Wald) tests. Specifically, I test the null hypothesis that the main effect and the interaction term are jointly equal to zero ($H_0: \beta_{Main} = \beta_{Interaction} = 0$). 


```{r joint-significance, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE, results='hold'}
library(car)

# custom function to execute and format robust joint significance tests
test_joint_significance <- function(model, var_main, var_interaction, test_name) {
  cat(paste0("====================================================\n"))
  cat(paste0("Joint Test: ", test_name, "\n"))
  cat(paste0("H0: ", var_main, " = 0 AND ", var_interaction, " = 0\n"))
  cat(paste0("====================================================\n"))
  
  # execute linear hypothesis test using clustered standard errors
  wald_result <- linearHypothesis(
    model, 
    c(paste0(var_main, " = 0"), paste0(var_interaction, " = 0")), 
    vcov. = vcovHC(model, type = "HC1", cluster = "group")
  )
  
  print(wald_result)
  cat("\n\n")
}

# quality joint tests
cat("### PATENT QUALITY MODELS ###\n\n")

# market Finance Model
test_joint_significance(market_quality[[4]], "ln_market_c", "ln_market_x_ipr", "Market Cap & Interaction (Quality)")
test_joint_significance(market_quality[[4]], "ipr_c", "ln_market_x_ipr", "IPR & Interaction [Market Model] (Quality)")

# Bank Credit Model
test_joint_significance(bank_quality[[4]], "ln_bank_c", "ln_bank_x_ipr", "Bank Credit & Interaction (Quality)")
test_joint_significance(bank_quality[[4]], "ipr_c", "ln_bank_x_ipr", "IPR & Interaction [Bank Model] (Quality)")


# patent quanity joint test
cat("### PATENT QUANTITY MODELS ###\n\n")

# Market Finance Model
test_joint_significance(market_quantity[[4]], "ln_market_c", "ln_market_x_ipr", "Market Cap & Interaction (Quantity)")
test_joint_significance(market_quantity[[4]], "ipr_c", "ln_market_x_ipr", "IPR & Interaction [Market Model] (Quantity)")

# Bank Credit Model
test_joint_significance(bank_quantity[[4]], "ln_bank_c", "ln_bank_x_ipr", "Bank Credit & Interaction (Quantity)")
test_joint_significance(bank_quantity[[4]], "ipr_c", "ln_bank_x_ipr", "IPR & Interaction [Bank Model] (Quantity)")
```
## Country-specific time trends

```{r time-trends, results='asis', echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}
library(data.table)
library(dplyr)
library(fixest) # replaces plm, lmtest, and sandwich
# it was harder to do this with plm so i switched packages

options(dplyr.summarise.inform = FALSE)

run_trend_incremental <- function(depvar, fd, fd_x_ipr, title) {

  # maximize samples by finance type
  base <- df %>%
    filter(
      !is.na(.data[[depvar]]),
      !is.na(.data[[fd]]),
      !is.na(ipr_c)
    )

  # ensure year is numeric so fixest treats iso3c[year_numeric] as a linear trend
  base$year_numeric <- as.numeric(as.character(base$year))
  
  # specify the fixed effects and trend part of the formula
  fe_trend <- "| iso3c + year + iso3c[year_numeric]"
  
  # build specifications
  f1 <- as.formula(paste(depvar, "~", fd, "+ ipr_c +", fd_x_ipr, fe_trend))
  f2 <- as.formula(paste(depvar, "~", fd, "+ ipr_c +", fd_x_ipr, "+ ln_gdp_pc", fe_trend))
  f3 <- as.formula(paste(depvar, "~", fd, "+ ipr_c +", fd_x_ipr, "+ ln_gdp_pc + ln_trade", fe_trend))
  f4 <- as.formula(paste(depvar, "~", fd, "+ ipr_c +", fd_x_ipr, "+ ln_gdp_pc + ln_trade + ln_rd", fe_trend))
  f5 <- as.formula(paste(depvar, "~", fd, "+ ipr_c +", fd_x_ipr, "+ ln_gdp_pc + ln_trade + ln_rd + ln_tertiary", fe_trend))

  # run models with fixest (cluster = ~iso3c applies HC1 clustered standard errors)
  m1 <- feols(f1, data = base, cluster = ~iso3c)
  m2 <- feols(f2, data = base, cluster = ~iso3c)
  m3 <- feols(f3, data = base, cluster = ~iso3c)
  m4 <- feols(f4, data = base, cluster = ~iso3c)
  m5 <- feols(f5, data = base, cluster = ~iso3c)
  
  # format custom labels for the table
  var_dict <- c(
    quality_index_100 = "Patent Quality (0–100)", 
    ln_patents_pc = "Log Patents per Capita",
    ln_market_c = "Log Market Cap (Centered)",
    ln_bank_c = "Log Bank Credit (Centered)",
    ln_vc_c = "Log Venture Capital (Centered)",
    ipr_c = "IPR Index (Centered)",
    ln_market_x_ipr = "Log Market Cap x IPR",
    ln_bank_x_ipr = "Log Bank Credit x IPR",
    ln_vc_x_ipr = "Log Venture Capital x IPR",
    ln_gdp_pc = "Log GDP per Capita",
    ln_trade = "Log Trade Openness",
    ln_rd = "Log R&D Expenditure",
    ln_tertiary = "Log Tertiary Enrollment"
  )


  cat(paste0("### ", title, "\n\n")) 
  

  table_text <- capture.output(
    etable(
      m1, m2, m3, m4, m5,
      dict = var_dict,
      signif.code = c("***"=0.01, "**"=0.05, "*"=0.1),
      fitstat = ~ n + r2
    )
  )
  

  cat("```text\n")
  cat(paste(table_text, collapse = "\n"))
  cat("\n```\n")
  

  cat("\n*Notes: Country and year fixed effects included. Country-specific linear time trends included. Standard errors clustered by country.*\n")
  cat("<br><br>\n\n") 

  invisible(list(m1, m2, m3, m4, m5))
}


# quality models-

market_quality <- run_trend_incremental(
  depvar   = "quality_index_100",
  fd       = "ln_market_c",       
  fd_x_ipr = "ln_market_x_ipr",   
  title    = "Market Finance and Patent Quality (With Time Trends)"
)

bank_quality <- run_trend_incremental(
  depvar   = "quality_index_100",
  fd       = "ln_bank_c",         
  fd_x_ipr = "ln_bank_x_ipr",     
  title    = "Bank Finance and Patent Quality (With Time Trends)"
)

vc_quality <- run_trend_incremental(
  depvar   = "quality_index_100",
  fd       = "ln_vc_c",           
  fd_x_ipr = "ln_vc_x_ipr",       
  title    = "Venture Capital and Patent Quality (With Time Trends)"
)

# quanitity models

market_quantity <- run_trend_incremental(
  depvar   = "ln_patents_pc",
  fd       = "ln_market_c",       
  fd_x_ipr = "ln_market_x_ipr",
  title    = "Market Finance and Patent Quantity (With Time Trends)"
)

bank_quantity <- run_trend_incremental(
  depvar   = "ln_patents_pc",
  fd       = "ln_bank_c",         
  fd_x_ipr = "ln_bank_x_ipr",
  title    = "Bank Finance and Patent Quantity (With Time Trends)"
)

vc_quantity <- run_trend_incremental(
  depvar   = "ln_patents_pc",
  fd       = "ln_vc_c",           
  fd_x_ipr = "ln_vc_x_ipr",
  title    = "Venture Capital and Patent Quantity (With Time Trends)"
)
```

# Variable Glossary
**[Source for PQI Data](https://forms.office.com/pages/responsepage.aspx?id=1MdBrGEfDUaw9PySWitHHKuxmuqpz_9KusL7-G1D6wFUOEU0OVBYVk5QTzROVlBTSUtBUUREWVhHTiQlQCN0PWcu&route=shorturl)** 
**[Source for WDI data](https://databank.worldbank.org/source/world-development-indicators)**
**[Source for VC data](https://data-explorer.oecd.org/vis?df[ds]=DisseminateFinalDMZ&df[id]=DSD_VC%40DF_VC_INV&df[ag]=OECD.SDD.TPS&dq=...USD_EXC.A&pd=2007%2C&to[TIME_PERIOD]=false)**
**[Source for Park Index](https://edspace.american.edu/waltergpark/2025/09/01/patent-rights-index/)**

### Dependent Variables
- `quality_index_100` - The OECD composite patent quality index (`avg_quality_index_4`), scaled to a 0–100 range for interpretability.
- `patents_per_capita` - Patent applications by residents, scaled per million people.
- `ln_patents_pc` - Log-transformed patenting intensity ($\log(x+1)$ of `patents_per_capita`). This replaces previous researcher-scaled metrics to better capture output while maximizing sample size.
- `breakthrough_count` - The absolute count of breakthrough patents (top 1% most cited) produced by a country in a given year. Used exclusively as the dependent variable in the ZINB models.

### Financial Development Variables (Independent)
- `bank_credit_gdp` - Domestic credit provided by the financial sector (% of GDP). Proxy for traditional banking depth.
- `market_cap_gdp` - Market capitalization of listed domestic companies (% of GDP). Proxy for public equity market depth.
- `VC_investment_pct` - Venture capital investment as % of GDP (rescaled by $\times 100$ to match the percentage units of the other financial variables).
- `ln_bank`, `ln_market`, `ln_vc` - Log-transformed ($\log(x+1)$) versions of the respective financial metrics to address right-skewness.
- `ln_bank_c`, `ln_market_c`, `ln_vc_c` - Mean-centered versions of the log-transformed financial variables. Centering ensures that the main effects in the interaction models are evaluated at the sample mean.

### Institutional Variables (Independent)
- `ipp_index_IP_transformed` - The Ginarte-Park Intellectual Property Protection Index (0–5 scale), using a standard five-year fill interpolation for missing inter-period years.
- `ipr_c` - Mean-centered Park Index.

### Interaction Terms
- `ln_bank_x_ipr`, `ln_market_x_ipr`, `ln_vc_x_ipr` - The interaction between the mean-centered financial variable and the mean-centered Park Index ( `ln_market_c` $\times$ `ipr_c`). 

### Macroeconomic Controls
*Note: All continuous macro controls are $\log(x+1)$ transformed to capture proportional relationships and minimize the influence of extreme outliers.*
- `ln_gdp_pc` - Log of real GDP per capita (constant 2015 USD).
- `ln_trade` - Log of trade openness (Imports + Exports as % of GDP).
- `ln_rd` - Log of Gross Domestic Expenditure on R&D (% of GDP).
- `ln_tertiary` - Log of gross tertiary school enrollment (%).
